---
title: Dashboard
author: Henry Linder (mhlinder@gmail.com)
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

# Setup

```{r message = FALSE}
library(magrittr)
library(dplyr)
library(knitr)
```

A cronjob executes every two weeks, which runs the file `get_data.R`
(the source is included at the end of this file). The script saves a
dated file in the directory `data/`.

```{r}
infiles <- list.files("data", full.names = TRUE)
n_infiles <- length(infiles)
infiles[n_infiles]
```

We load the most recent data file, which contains the master variable
`out`.

```{r}
load(infiles[n_infiles])

daily <-
    out$daily %>%
    as.data.frame %>%
    tbl_df

## Removes day zero, which does not record a full day
daily <- daily[-1,]
n <- nrow(daily)

## Removes day n, which is incomplete by dint of when the data was collected
daily <- daily[-n,]
n <- nrow(daily)

names(out)
```

# EDA

We start by considering daily historical step count, which is stored
in `daily$steps.time`. We also check whether there are any missing
days, because we assume there are none.

```{r fig.align = "center"}

dates <-
    daily$steps.time %>%
    as.Date

all_days <-
    seq.Date(dates[1], dates[n], by = "day") %>%
    as.character
d <- as.character(dates)

if (!all(d == all_days)) {
    stop("Missing days in input data!")
}

y <- daily$steps.steps

```

## Weekdays and weekends

We consider a simple time series plot to consider the data,

```{r fig.align = "center"}
## List of all weekday names, starts on Sunday
dows <- weekdays(x=as.Date(seq(7), origin="1949-12-31"))
wknd <- dows %in% c("Saturday", "Sunday")

obs_dow <- weekdays(dates) ## Weekday of each observation
is_wknd <- obs_dow %in% c("Saturday", "Sunday")

get_colors <- function(url = "https://coolors.co/033f63-28666e-7c9885-b5b682-fedc97") {
    ## `url` should be a link to a coolors.co color scheme
    ## Returns a vector with the palette's hex codes as character elements.
    basename(url) %>%
        strsplit("-") %>%
        unlist %>%
        paste0("#", .)
}

colors5 <- get_colors("https://coolors.co/033f63-28666e-7c9885-b5b682-fedc97")

colors7 <- c("#8569CF", "#0D9FD8", "#8AD749", "#EECE00",
             "#F8981F", "#F80E27", "#F640AE")
names(colors7) <- dows
tmp_color <- rep(colors5[1], 7)
tmp_color[wknd] <- "red"
names(tmp_color) <- dows

plot(dates, y, type = "l",
     xlab = "",
     ylab = "Steps",
     main = "Steps by day",
     xaxt = "n")
points(dates, y, col = tmp_color)

## Count backwards by week from the last observation
at <- seq.Date(dates[n], dates[1], by = "-1 week")
axis(1, at = at, labels = as.character(at))

abline(h = mean(y), lty = 2)
```

where the red dots label weekends, and the dotted line represents the
mean. Although the time series is noisy, there appears to be some
cyclical trend.

We also compare kernel density plots of the observations for weekends
and weekdays.

```{r fig.align = "center"}
par(mfrow = c(2,2))

plot(density(y[is_wknd]), main = "Weekends")

plot(density(y[!is_wknd]), main = "Weekdays")

qqnorm(y[is_wknd], main = "Weekends")
qqline(y[is_wknd])

qqnorm(y[!is_wknd], main = "Weekdays")
qqline(y[!is_wknd])
```

Notable features include the much longer right tail for the weekday
distribution, and the larger left tail for weekends. This reflects a
pattern that I walk the furthest on weekdays, sometimes quite far,
whereas on the weekends I am much less likely to go on a very long
walk.

Interestingly, though, the mean number of steps is higher for the
weekend than weekdays:

```{r}
mean(y[wknd])

mean(y[!wknd])
```

This can be understood at least partially as a n indication that there
is larger variability in how many steps I walk on weekdays, though on
average I walk fewer per day than on the weekends. And, of course, the
data reflects this:

```{r}
var(y[wknd])

var(y[!wknd])
```

## Day-of-the-week

Then, we can consider day-of-the-week breakdowns, like the average
steps per day:

```{r fig.align = "center"}
## Indexes of dates by day-of-week
ix_dow <- list()
for (d in dows) {
    ix_dow[[d]] <- which(obs_dow == d)
}

x <-
    ix_dow %>%
    lapply(function(ix) mean(y[ix])) %>%
    unlist

barplot(x, col = colors7,
        cex.names = .8,
        ylim = c(0, 10000), ylab = "Average steps")
```

Density plots for all days together, and by each day:

```{r fig.align = "center"}

par(mfrow = c(2, 4))

plot(density(y), main = "All days")
for (dow in dows) {
    plot(density(y[ix_dow[[dow]]]),
           main = dow, col = colors7[dow])
}
```

For all days, and each day, normal probability plots:

```{r fig.align = "center"}
par(mfrow = c(2, 4))

qqnorm(y, main = "All days")
qqline(y)
for (dow in dows) {
    qqnorm(y[ix_dow[[dow]]],
           main = dow, col = colors7[dow])
    qqline(y[ix_dow[[dow]]])
}
```

# Code

## `get_data.R`

```{r cache = FALSE, echo = FALSE}
read_chunk("get_data.R")
```
```{r get-data, eval = FALSE}
```

