---
title: Dashboard
author: Henry Linder (mhlinder@gmail.com)
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
    html_document:
        toc: true
        code_folding: hide
---

# Setup

```{r message = FALSE}
library(magrittr)
library(dplyr)
library(knitr)
```

## Graphical parameters

```{r}

get_colors <- function(url = "https://coolors.co/033f63-28666e-7c9885-b5b682-fedc97") {
    ## `url` should be a link to a coolors.co color scheme
    ## Returns a vector with the palette's hex codes as character elements.
    basename(url) %>%
        strsplit("-") %>%
        unlist %>%
        paste0("#", .)
}

colors5 <- get_colors("https://coolors.co/033f63-28666e-7c9885-b5b682-fedc97")
colors7 <- c("#8569CF", "#0D9FD8", "#8AD749", "#EECE00",
             "#F8981F", "#F80E27", "#F640AE")

## List of all weekday names, starts on Sunday
dows <- weekdays(x=as.Date(seq(7), origin="1949-12-31"))
wknd <- dows %in% c("Saturday", "Sunday")

names(colors7) <- dows
```

A cronjob executes every two weeks, which runs the file `get_data.R`
(the source is included at the end of this file). The script saves a
dated file in the directory `data/`.

```{r}
infiles <- list.files("data", full.names = TRUE)
n_infiles <- length(infiles)
infiles[n_infiles]
```

We load the most recent data file, which contains the master variable
`out`.

```{r}
load(infiles[n_infiles])

daily <-
    out$daily %>%
    as.data.frame %>%
    tbl_df %>%
    rename(steps = steps.steps)

## Removes day zero, which does not record a full day
daily <- daily[-1,]
n <- nrow(daily)

## Removes day n, which is incomplete by dint of when the data was collected
daily <- daily[-n,]
n <- nrow(daily)

names(out)
```

# EDA

We start by considering daily historical step count, which is stored
in `daily$steps.time`. We also check whether there are any missing
days, because we assume there are none.

```{r fig.align = "center"}

daily$dates <-
    daily$steps.time %>%
    as.Date

all_days <-
    seq.Date(daily$dates[1], daily$dates[n], by = "day") %>%
    as.character
d <- as.character(daily$dates)

if (!all(d == all_days)) {
    stop("Missing days in input data!")
}

```

## Weekdays and weekends

We consider a simple time series plot to consider the data,

```{r fig.align = "center"}
daily$dow <- weekdays(daily$dates) ## Weekday of each observation
daily$is_wknd <- daily$dow %in% c("Saturday", "Sunday")

## Find the first day, so the colors match the dates
st1 <- daily$dow[1]
ix_st <- which(colors7 %>% names == st1)
## This vector of indexes rearranges the week starting on day 1, to a
## week startin gon day ix_st
ix <- c((ix_st:7), (1:(ix_st-1)))

tmp_color <- rep(colors5[1], 7)
tmp_color[wknd] <- "red"
names(tmp_color) <- dows
tmp_color <- tmp_color[ix]

plot(daily$dates, daily$steps, type = "l",
     xlab = "",
     ylab = "Steps",
     main = "Steps by day",
     xaxt = "n")
points(daily$dates, daily$steps, col = tmp_color)

## Count backwards by week from the last observation
at <- seq.Date(daily$dates[n], daily$dates[1], by = "-1 week")
axis(1, at = at, labels = as.character(at))

abline(h = mean(daily$steps), lty = 2)
```

where the red dots label weekends, and the dotted line represents the
mean. Although the time series is noisy, there appears to be some
cyclical trend.

Further, on many occasions, the weekend points are very close to each
other---ie, their correlations seem high.

```{r}
sat <- daily %>% filter(dow == "Saturday")
sun <- daily %>% filter(dow == "Sunday")

if (nrow(sat) != nrow(sun)) {
    stop("Incomplete weekends!")
}

n_wknd <- nrow(sat)

ix_match <-
    sapply(1:n_wknd, function(x) {
        if (as.double(sun$dates[1] - sat$dates[1]) != 1) {
            FALSE
        } else {
            TRUE
        }})
sun <- sun[ix_match,]
sat <- sat[ix_match,]

cor(sat$steps, sun$steps)
```

This seems to suggest moderate association between the number of steps
taken on Saturdays and Sundays.

Several hypothesis tests provide compelling support that there is this
strong positive correlation between steps on these two days.

```{r}
cor.test(sat$steps, sun$steps, method = "pearson")
cor.test(sat$steps, sun$steps, method = "kendall")
cor.test(sat$steps, sun$steps, method = "spearman")
```

In other words, either it's a walking weekend, or a resting weekend,
but I rarely walk a lot on only one day of a weekend.

We also compare kernel density plots of the observations for weekends
and weekdays.

```{r fig.align = "center"}
par(mfrow = c(2,2))

data_weekends <-
    daily %>%
    filter(is_wknd)
data_weekdays <-
    daily %>%
    filter(!is_wknd)

plot(density(data_weekends$steps), main = "Weekends")

plot(density(data_weekdays$steps), main = "Weekdays")

qqnorm(data_weekends$steps, main = "Weekends")
qqline(data_weekends$steps)

qqnorm(data_weekdays$steps, main = "Weekdays")
qqline(data_weekdays$steps)
```

Notable features include the much longer right tail for the weekday
distribution, and the larger left tail for weekends. This reflects a
pattern that I walk the furthest on weekdays, sometimes quite far,
whereas on the weekends I am much less likely to go on a very long
walk.

Interestingly, though, the mean number of steps is higher for the
weekend than weekdays:

```{r}
mean(data_weekends$steps)

mean(data_weekdays$steps)
```

This can be understood at least partially as a n indication that there
is larger variability in how many steps I walk on weekdays, though on
average I walk fewer per day than on the weekends. And, of course, the
data reflects this:

```{r}
sd(data_weekends$steps)

sd(data_weekdays$steps)
```

## Day-of-the-week

Then, we can consider day-of-the-week breakdowns, like the average
steps per day:

```{r fig.align = "center"}
## Indexes of dates by day-of-week
ix_dow <- list()
for (d in dows) {
    ix_dow[[d]] <- which(daily$dow == d)
}

x <-
    ix_dow %>%
    lapply(function(ix) mean(daily$steps[ix])) %>%
    unlist

barplot(x, col = colors7,
        cex.names = .8,
        ylim = c(0, 10000), ylab = "Average steps")
```

These daily values look pretty similar to each other, and an ANOVA
test for a difference in means may provide a useful way to summarize
the data.

Density plots for all days together, and by each day:

```{r fig.align = "center"}

par(mfrow = c(2, 4))

plot(density(daily$steps), main = "All days")
for (dow in dows) {
    plot(density(daily$steps[ix_dow[[dow]]]),
           main = dow, col = colors7[dow])
}
```

For all days, and each day, normal probability plots:

```{r fig.align = "center"}
par(mfrow = c(2, 4))

qqnorm(daily$steps, main = "All days")
qqline(daily$steps)
for (dow in dows) {
    qqnorm(daily$steps[ix_dow[[dow]]],
           main = dow, col = colors7[dow])
    qqline(daily$steps[ix_dow[[dow]]])
}
```

These distributions are symmetric and approximately unimodal, so the
data can be said to be approximately normal---at least for our
purposes, "approximate" normality suffices. This is because ANOVA is
robust to deviations from normality, generally speaking.

Recall from above that the weekend and weekday datasets were each
approximately normal, so we can conduct two ANOVA tests.

```{r}
aov(steps ~ dow, data = daily) %>%
    summary
```

The first ANOVA considers each day to be a separate population, and
the test pretty conclusively fails to reject the null hypothesis of
zero difference. In other words, **from a statistical perspect, there
is no difference in average numbers of steps per day between
different days of the week.**

```{r}
aov(steps ~ is_wknd, data = daily) %>%
    summary
```

The test-statistic is larger, the p-value is smaller, but there is
still effectively zero support for the hypothesis that my walking
habits are different on the weekends from the weekdays, at least in
terms of the number of steps I take.

In other words, it does not seem meaningful to consider each day of
the week separately, or weekdays and weekends separately. There are
likely other, more nuanced subsets possible for the days of the weeK:
for instance, last fall I took courses on Tuesdays and Thursdays, and
on those days I had a long walk to and from my car.

```{r}
daily$ttr <- daily$dow %in% c("Tuesday", "Thursday")
aov(steps ~ ttr, data = daily) %>%
    summary
```

This doesn't show up, either.


# Code

## `get_data.R`

```{r cache = FALSE, echo = FALSE}
read_chunk("get_data.R")
```
```{r get-data, eval = FALSE}
```

